---
title: "Assignment 7 - Repeated measure"
author: "Aytan Pirverdiyeva"  
format:
  html:
    code-fold: true 
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Introduction  
This assignment will ask you to conduct a repeated measure analysis. 

For that, you will first need to watch (and code along with) the following repeated measure video:

- repeated measure lecture: https://youtu.be/-ANqcQhB--A?si=7YeeYqDgcLsj5erW&t=4268 (starts at 01:11:08, watch till the end, about 50 min)

- repeated measure lab part 1: https://youtu.be/GY2e-9M5bFM?si=TIw5F4EF2m9Dx5uK (watch full video, ~1h 50 min)  

- repeated measure lab part 2: https://youtu.be/F3vg89ymLeM?si=Zv-RPIdS8mig4EMb
(watch full video, ~2h)

- The partial script for the video above is in our GitHub repository, 04-classcode, **03-11_repeatedmeasure_partial.qmd**.  

- The data set for this exercise is in oir GitHub repository, 03-classcode, **wheat_nk_balkh_rm.csv**.


**AFTER watching the video and developing the code in class**, then start working on this HW assignment.  

# Instructions  

  - Create on your GitHub a repo named `06_rm`.  
  - Clone it to your computer, create the necessary subfolders.
  - Move the script **03-11_repeatedmeasure_partial.qmd** into the `code` subfolder.
  - Move this script into the `code` subfolder  
  - Move the **wheat_nk_balkh_rm.csv** data set into the `data` subfolder  
  - On this script, fill your name under the `title` section of the YAML  
  
# Goal of this assignment  
The goal is this assignment is for you to **learn about and conduct a full repeated measure analysis**.

Once you finish coding along with the videos, answer the following questions.

#Setup and Wrangling
```{r}
#| message: false
#| warning: false
library(tidyverse) # for data wrangling and plotting
library(car) # for Anova function
library(lme4)
library(broom) # for model residuals extraction
library(emmeans) # for model mean extraction
library(multcomp) # for pairwise comparison letter display
library(nlme) # for repeated measure correlation structures 
```

#Read data
```{r}
rm_rcbd_df <- read_csv("../data/wheat_nk_balkh_rm.csv") %>%
  mutate(across(trt:splot, ~factor(.))) %>%
  mutate(fdap = factor(dap))


rm_rcbd_df
```

#EDA
```{r}
summary(rm_rcbd_df)

```
```{r}
ggplot(rm_rcbd_df, aes(x = fdap, 
                       y = yield_kgha))+
  geom_boxplot() + 
  facet_grid(.~trtname)
```


#Model 1 - Default
```{r}
options(contrasts = c("contr.sum", "contr.poly"))

mod1_default_lmer <-lmer(yield_kgha ~ krate_kgha*nrate_kgha*fdap + 
                           (1|block/krate_kgha/nrate_kgha),
                         data = rm_rcbd_df
                           )
mod1_default_lmer
```


```{r}
mod1_default <-lme(yield_kgha ~krate_kgha*nrate_kgha*fdap,
                   random = ~1 |block/krate_kgha/nrate_kgha,
                   data = rm_rcbd_df
                   )


mod1_default
```

#Check Anova
```{r}
Anova(mod1_default, type = 3)
```


```{r}
ACF(mod1_default, resType = "n") %>%
  plot(alpha = 0.01)

```

# Model 2 - Compound Symmetry  
```{r}
mod2_cs <- lme(yield_kgha ~krate_kgha*nrate_kgha*fdap,
                   random = ~1 |block/krate_kgha/nrate_kgha,
                   correlation = corCompSymm(form = ~dap),
                   data = rm_rcbd_df
                   )

mod2_cs
```

```{r}
ACF(mod2_cs, resType = "n") %>%
  plot(alpha = 0.01)
```


```{r}
intervals(mod2_cs,which = "var-cov")
```

```{r}
anova(mod1_default,
      mod2_cs)
```

# Questions  
In class, we decided to use AIC as the metric to select the best model, and this determined the best model to be the one with a general error correlation matrix structure. If we had used BIC as the metric, then the model selected would have been the default model. Here, you are asked to use BIC and thus select the default model as the best, and perform all the steps after model selection. You may copy and paste the class script, and adapt the copied version to implement these changes.  

## Q1: Include here the residual assessment of the default model. Include each plot and its interpretation. How do these residuals from the default model visually compare to those from the general model?

```{r}
mod1_default <-lme(yield_kgha ~krate_kgha*nrate_kgha*fdap,
                   random = ~1 |block/krate_kgha/nrate_kgha,
                   data = rm_rcbd_df
                   )


mod1_default

```


```{r}
par(mfrow = c(2, 2))
plot(mod1_default, which = 1)  # Residuals vs Fitted

qqnorm(resid(mod1_default)); qqline(resid(mod1_default))  # Normality check

hist(resid(mod1_default), main = "Histogram of Residuals")

acf(resid(mod1_default), main = "ACF of Residuals")

```
#Interpretation
While the general model may exhibit slightly better residual behavior, the default model's residuals were acceptable and satisfied the main assumptions, making it suitable for analysis based on BIC selection. 
The residual assessment of the default model with compound symmetry revealed that the model fits the data reasonably well; the Residuals vs. Fitted plot showed a random scatter without any clear patterns, indicating constant variance; the QQ plot showed that most residuals closely followed the straight line, suggesting the normality assumption was reasonably met; the residuals histogram seemed bell-shaped and symmetrical, further supporting a normal distribution; the ACF plot showed no significant autocorrelation, indicating the model properly accounted for the repeated measures structure; and the default model's residuals were acceptable and satisfied the main assumptions.

## Q2: Include here the ANOVA table from the default model. How does it compare with the ANOVA table from the general model?
```{r}
anova(mod1_default)
```
#Interpretation
Although the main time effect (fdap) and the potassium-nitrogen interaction were both significant in the default model, the general model probably had slightly higher significance levels, especially for complex interactions, because it better explained the correlation in repeated measures. For instance, the three-way interaction (krate_kgha:nrate_kgha:fdap) was only marginally significant (p = 0.0926) in the default model but may be more significant in the general model.

# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE` (I'll demo this in a moment).  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one (I'lld demo this in a moment).  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `Assignment-07-rm-LASTNAME.html`.    

For ex., mine would be `Assignment-07-rm-Bastos.html`.

Submit your work by uploading the **html** file to **eLC Assigment #7 - Repeated measure** by Mar 25 11:59 pm.  
  




